---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# slimr

[https://rdinnager.github.io/slimr/](https://rdinnager.github.io/slimr/)

<!-- badges: start -->
[![R build status](https://github.com/rdinnager/slimr/workflows/R-CMD-check/badge.svg)](https://github.com/rdinnager/slimr/actions)
[![Codecov test coverage](https://codecov.io/gh/rdinnager/slimr/branch/master/graph/badge.svg)](https://codecov.io/gh/rdinnager/slimr?branch=master)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of slimr is to run SLiM population genetics forward simulations from R. It also has utilities for monitoring the simulations, and bringing the resulting data into R for post-processing and visualization.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("rdinnager/slimr")
```

## Writing SLiM Scripts in R

Using `slimr`, this is how you write the first example script (or recipe) from the (excellent) [SLiM](https://messerlab.org/slim/) manual:

```{r recipe_1}
library(slimr)

slim_script(
  slim_block(initialize(),
             {
               ## set the overall mutation rate
               initializeMutationRate(1e-7); 
               ## m1 mutation type: neutral
               initializeMutationType("m1", 0.5, "f", 0.0);
               ## g1 genomic element type: uses m1 for all mutations
               initializeGenomicElementType("g1", m1, 1.0);
               ## uniform chromosome of length 100 kb
               initializeGenomicElement(g1, 0, 99999);
               ## uniform recombination along the chromosome
               initializeRecombinationRate(1e-8);
             }),
  slim_block(1,
             {
               sim.addSubpop("p1", 500);
             }),
  slim_block(10000,
             {
               sim.simulationFinished();
             })
) -> script_1

script_1
```

You can output this script to text to run in a standalone SLiM installation, or you can run it in SLiM directly from R.

You can also do fancy stuff like make the above script or another script into a template that you can dynamically fill-in with parameters generated in R. You can also make SLiM generate R-friendly input. See the vignettes for details of these features and how to use them.

