---
title: "How to Install SLiM for Use with slimr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{installing_SLiM_for_slimr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(slimr, quietly = TRUE)
```

This vignette will try and guide you through installing SLiM on your machine to use with `slimr`.

After installing `slimr` and loading it for the first time you will probably see the following message:

```{r message, echo=FALSE}
message("Welcome to the slimr package. slimr requires SLiM population genetics forward simulation software. Run slim_setup() for slimr to attempt to download and install the package automatically. If that doesn't work, try installing manually by going to https://messerlab.org/slim/ and following the instructions.")
```

This let's you know that SLiM is not yet installed, and you will need to install it in order to use much of the functionality of `slimr`. It suggests trying the function `slim_setup()`. It should work on any platform, though it has only been tested extensively on Windows and Linux. Since SLiM is only available for unix based systems, in order to use SLiM on Windows, you will need to have Windows Subsystem for Linux (WSL) installed on your system. WSL is available for Windows 10. If it is not already installed, you can follow the directions here to install it: [https://docs.microsoft.com/en-us/windows/wsl/install-win10](https://docs.microsoft.com/en-us/windows/wsl/install-win10). We recommend that you install the newest Ubuntu flavour when choosing a linux distribution, as `slimr` has only been tested on Ubuntu WSL. 
