---
title: "simple_example_using_migration_and_fst"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r color, echo = FALSE, results='asis'}
# crayon needs to be explicitly activated in Rmd
options(crayon.enabled = TRUE)
# Hooks needs to be set to deal with outputs
# thanks to fansi logic
if(requireNamespace("fansi", quietly = TRUE)) {
  old_hooks <- fansi::set_knit_hooks(knitr::knit_hooks, 
                                     which = c("output", "message", "error"))
}
```


```{r setup}
library(slimr)
library(adegenet)
library(dartR)
library(ggplot2)
library(purrr)
library(furrr)
set.seed(123456)
```


```
initialize() {
  initializeMutationRate(1e-7);
  initializeMutationType("m1", 0.5, "f", 0.0);
  initializeGenomicElementType("g1", m1, 1.0);
  initializeGenomicElement(g1, 0, 99999);
  initializeRecombinationRate(1e-8);
}

1 {
  for (i in 1:3)
    sim.addSubpop(i, 1000);
  subpops = sim.subpopulations;
  lines = readFile("~/Desktop/migration.csv");
  lines = lines[substr(lines, 0, 1) != "//"];
  for (line in lines)
  {
    fields = strsplit(line, ",");
    i = asInteger(fields[0]);
    j = asInteger(fields[1]);
    m = asFloat(fields[2]);
    if (i != j)
    {
      p_i = subpops[subpops.id == i];
      p_j = subpops[subpops.id == j];
      p_j.setMigrationRates(p_i, m);
    }
  }
}

10000 late() { sim.outputFull(); }

```

```{r script}
slim_script(
  slim_block(initialize(), {
    initializeMutationRate(1e-7);
    initializeMutationType("m1", 0.5, "f", 0.0);
    initializeGenomicElementType("g1", m1, 1.0);
    initializeGenomicElement(g1, 0, 99999);
    initializeRecombinationRate(1e-8);
  }),
  slim_block(1, {
    for (i in 1:3) {
      sim.addSubpop(i, 1000);
    }
    subpops = sim.subpopulations;
    lines = readFile("~/Desktop/migration.csv");
    lines = lines[substr(lines, 0, 1) != "//"];
    for (line in lines) {
      fields = strsplit(line, ",");
      i = asInteger(fields[0]);
      j = asInteger(fields[1]);
      m = asFloat(fields[2]);
      if (i != j) {
        p_i = subpops[subpops.id == i];
        p_j = subpops[subpops.id == j];
        p_j.setMigrationRates(p_i, m);
      }
    }
  }),
  slim_block(10000, late(), {
    sim.outputFull();
  })
) -> script_1

script_1

```

```{r script_2}

disp_mat <- matrix(c(1, 1, 0.78, 
                     1, 2, 0.10, 
                     1, 3, 0.12, 
                     2, 1, 0.01, 
                     2, 2, 0.96, 
                     2, 3, 0.03, 
                     3, 1, 0.33, 
                     3, 2, 0.17, 
                     3, 3, 0.5),
                   ncol = 3, byrow = TRUE)

slim_script(
  slim_block(initialize(), {
    setSeed(123456);
    initializeMutationRate(2e-6);
    initializeMutationType("m1", 0.5, "f", 0.0);
    initializeGenomicElementType("g1", m1, 1.0);
    initializeGenomicElement(g1, 0, 99999);
    initializeRecombinationRate(1e-8);
  }),
  slim_block(1, {
    for (i in 1:3) {
      sim.addSubpop(i, 100);
    }
    subpops = sim.subpopulations;
    
    disp_mat = slimr_inline(disp_mat)
    
    for (line in seqLen(nrow(disp_mat))) {
      i = asInteger(disp_mat[line, 0]);
      j = asInteger(disp_mat[line, 1]);
      m = asFloat(disp_mat[line, 2]);
      if (i != j) {
        p_i = subpops[subpops.id == i];
        p_j = subpops[subpops.id == j];
        p_j.setMigrationRates(p_i, m);
      }
    }
  }),
  slim_block(100000, late(), {
    slimr_output(sim.outputFull(), "final_output");
  })
) -> script_2

script_2

```
We can see that the whole matrix has been embedded in the SLiM script by running `as_slim_text` on this `slimr_script` object. This is the script that will actually be run by SLiM.

```{r show_script}
cat(as_slim_text(script_2))
```

Now let's see if we can run it.

```{r run_script}
tfile <- "temp.txt"
results <- slim_run(script_2, capture_output = tfile, progress = FALSE)

```

Looks like it worked. We can get the resulting genomic data from the returned object and convert it into a genlight object using the `slim_output_genlight()` function. Then we can plot the results.

```{r get_dat}
library(dartR)

inds <- slim_output_genlight(results, "final_output")
plot(inds)

```
Note that the SNPs are ordered by their mutation id, which should roughly correspond to their order of origination, so the leftmost mutations are the oldest, the rightmost the youngest. Using a genlight object we can calculate a lot of population genetics metrics easily using the package `dartR`. Let's calculate Fst for example.

```{r pop_gen}
p_fst <- dartR::gl.fst.pop(inds, nboots = 1)
p_fst
```
So generally those Fst values are low, which we would expect since our populations are not experiencing differential selection and the effects of drift are counteracted by migration between the populations. Nevertheless the pairwise Fst's measured here are inversely proportional to the overall strength of migration between the population in our simulation. We can see this like this:

```{r compare_mig}
## convert migration rates into pairwise matrix using igraph
library(igraph)
overall_mig_rates <- disp_mat %>%
  as.data.frame() %>%
  igraph::graph_from_data_frame() %>%
  igraph::as.undirected(mode = "collapse", edge.attr.comb = "mean") %>%
  igraph::as_adjacency_matrix(type = "lower", attr = "V3", sparse = FALSE)

diag(overall_mig_rates) <- NA
overall_mig_rates[upper.tri(overall_mig_rates)] <- NA

p_fst
overall_mig_rates

plot(as.vector(overall_mig_rates), as.vector(p_fst))
```

We can see generally the populations with the lowest migration rates also has the highest pairwise Fst, suggesting population with low migration have differentiated the most, which makes sense. In this particular example, the Fst seems to reach a minimum soon after, such that the two higher migration pairs don't differ much in their pairwise Fsts. This could be a result of stochasticity. if we really want to know whether migration rate effects Fst under a simple neutral model of evolution, we should run many simulations and look at the average effect. `slimr` makes this easy to do using templating. Let's simplify this model a little to just look at two populations at a time, and then set it up to run many times with diiferent migration rates between them. Ww can also vary population size as it seems like this should be an important mediator of any effects, given it will effect the relative strength of drift in the populations.

```{r templated}

slim_script(
  slim_block(initialize(), {
    initializeMutationRate(2e-6);
    initializeMutationType("m1", 0.5, "f", 0.0);
    initializeGenomicElementType("g1", m1, 1.0);
    initializeGenomicElement(g1, 0, 99999);
    initializeRecombinationRate(1e-8);
  }),
  slim_block(1, {
    sim.addSubpop("p1", slimr_template("pop_size", default = 100));
    sim.addSubpop("p2", ..pop_size..);
    p1.setMigrationRates(p2, slimr_template("m", default = 0.1));
    p2.setMigrationRates(p1, ..m..);
  }),
  slim_block(10000, late(), {
    slimr_output(sim.outputFull(), "final_output");
  })
) -> script_3

script_3

```

So in the above script we now have a placemarker for a `pop_size` and a `m` parameter. We specified this using the `slimr_template` function directly inserted into the SLiM code. We also demonstrate a trick to reduce typing. We only have to use the `slimr_template` function once for a given variable, subsequent references to the same variable can be specified directly using the `..x..` notation. Now to generate a script with the templated variables replaced with values of our choosing, we use the `slim_script_render` function. Here is an example:

```{r slimr_template_example}

example_script <- slimr_script_render(script_3, template = list(pop_size = 50, m = 0.05))
example_script

```

We can also pass a `data.frame` to `slim_script_render` which will allow us to generate many scripts at once, let's setup a large number of scripts to explore different values of our two parameters.

```{r setup_many_scripts}

param_df <- data.frame(pop_size = sample(5:100, 1000, replace = TRUE),
                                                         m = runif(1000, 0, 0.5))

script_list <- slimr_script_render(script_3,
                                   template = param_df)

script_list

```

This function returns a `slimr_script_coll` object, or a slimr script collection. We can now use `slim_run` on this object, which will run every script in the collection, one after the other, and collect the results. We can specify a `parallel = TRUE` argument to run the scripts in parallel. `slimr` uses the `future` package for parallelization. This means that the user must specify a future `plan` prior to running `slim_run` which will setup the workers to be used by `slimr`. If you want to use parallelization with `slimr`, we recommend you familiarize yourself with the `future` package. If in doubt `future::plan(future::multisession)` is the most general purpose specification which should work well on all platforms (and is the recommended plan by future for best stability of parallelization). You must also ensure the `future` and `furrr` packages are installed for this functionality to work. Let's try running these:

```{r run_all_scripts}
start_time <- Sys.time()
if(requireNamespace("furrr", quietly = TRUE)) {
  
  future::plan(future::multisession)
  
  all_results <- slim_run(script_list, parallel = TRUE, progress = TRUE)
  
}


```

Cool! First let's check if all simulation completed without error:

```{r check_errors}

errors <- purrr::map_lgl(all_results,
                         ~length(.x$error) > 0)
sum(errors)

```

Let's extract all the genomic data into genlight objects so we can easily calculate Fst. Since this might also take a little while we will parallelize this extraction as well using `furrr` (since we already set a `future` plan, and `furrr` uses `future` under the hood). Just in case any simulations give us trouble, we will wrap the extraction in `try()` so that the whole thing doesn't fail for one error.

```{r all_genlight}
all_gl <- furrr::future_map(all_results,
                     ~suppressMessages(try(slim_output_genlight(.x, "final_output"))),
                     .progress = TRUE)
```

Any errors?

```{r check_errors2}
errored <- which(sapply(all_gl, function(x) class(x) == "try-error"))
errored

purrr::map(errored,
           ~all_results[[.x]]$output_data$data)

```

We got 'r length(errored)' errors. Examining their data strings, we can see that they all essentially had no mutations, which is why we couldn't construct a genlight object.
Let's just filter these out and continue with our analysis by calculating Fst.

```{r get_fsts}
if(length(errored) > 0) {
  all_gl <- all_gl[-errored]
}

all_fsts <- furrr::future_map_dbl(all_gl,
                                  ~dartR::gl.fst.pop(.x, nboots = 1)[2, 1],
                                  .progress = TRUE)

end_time <- Sys.time()

end_time - start_time

cat("Total Duration:", end_time - start_time)
```

Firstly, negative values of Fst are incorrect, and usually reflect numerical inaccuracy around zero, so we will set negative Fst values to zero. Then let's look at the distribution.

```{r fst_dist}
all_fsts[all_fsts < 0] <- 0.0
hist(all_fsts, breaks = 100)

```
So how is Fst related to migration rate and population size? We can get an idea by running a simple linear model. Then we will do plots.

```{r fst_model}

if(length(errored) > 0) {
  param_df <- param_df[-errored, ]
}

fst_dat <- param_df %>%
  dplyr::mutate(fst = all_fsts,
                pop_size_st = (pop_size - mean(pop_size)) / sd(pop_size),
                m_st = (m - mean(m)) / sd(m))

mod <- lm(fst ~ pop_size_st*m_st, data = fst_dat)
summary(mod)

```

So both migration rate

We can demonstrate the interaction by splitting our data into population size bins and plotting the migration rate relationship as separate lines for each bin.

```{r plot_that}
fst_dat <- fst_dat %>%
  dplyr::mutate(ps_bin = cut(pop_size, 5, ordered_result = TRUE))

ggplot(fst_dat, aes(m, fst)) +
  geom_point(aes(colour = ps_bin)) +
  geom_smooth(aes(colour = ps_bin, fill = ps_bin)) +
  scale_x_sqrt() +
  theme_minimal()

```
